\documentclass[a4paper,12pt,oneside]{book}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{multirow}
\usepackage{url}
\usepackage[ruled,vlined, linesnumbered]{algorithm2e}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{example}{Example}[section]
\numberwithin{equation}{section}
\usepackage{changepage}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newtheorem{oss}{Remark}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\setlength\parindent{0pt}
\usepackage{pgfplots}
\usepackage{tikz}
\usepackage{caption}
\usepackage{subfig}
\tikzset{
  font={\fontsize{9pt}{12}\selectfont}}
\SetKwRepeat{Do}{do}{while}
\renewcommand{\arraystretch}{2}
\title{Optimization Methods Applied to Portfolio Selection}
\author{Federico D'Amato}
 
\begin{document}
 
\maketitle\clearpage
\tableofcontents\clearpage

\chapter{Introduction}
\input{introduction}


\chapter{Risk Parity portfolio}
\input{riskparity}


\chapter{Algorithms for computing portfolios}
\input{algorithms}

\chapter{A decomposition framework}
\input{decomposition}


\chapter{Computational Experiments}
\input{experiments}

\chapter{Conclusions}
\chapter{Acknowledgements}
\clearpage


\begin{thebibliography}{9}
\bibitem{markovitz}
  H. Marling, S. Emanuelsson,
  \emph{The Markowitz Portfolio Theory},
  2012.
\bibitem{sharpe}
  W. F. Sharpe,
  \emph{The Sharpe Ratio},
  1994.
\bibitem{diversification}
  Y. Choueifaty, T. Froidure, J. Reynier,
  \emph{Properties of the most diversified portfolio},
  2013. 
\bibitem{libro}
  G. Cornuejols, R. Tutuncu,
  \emph{Optimization methods in finance}, pp.138-140,
   2007.
\bibitem{tutuncu}
  X. Bai, K. Scheinberg, R. Tutuncu,
  \emph{Least-square approach to risk parity in portfolio selection},
  2013.   
\bibitem{bruder}
  B. Bruder, T. Roncalli,
  \emph{Managing Risk Exposures using the Risk Budgeting Approach},
  2012.   
\bibitem{roncalli}
  T. Roncalli,
  \emph{From portfolio optimization to risk parity},
  2012.
\bibitem{intr}
  T.Roncalli,
  \emph{Introducing Expected Returns into Risk Parity Portfolios: A New Framework for Asset Allocation},
  2014.   
\bibitem{erc}
  S. Maillard, T. Roncalli, J. Teiletche,
  \emph{On the properties of equally-weighted risk contributions portfolios},
  2009.
\bibitem{colucci}
  F. Cesarone, S. Colucci,
  \emph{Minimum Risk vs. Capital and Risk Diversification strategies for portfolio construction},
  2015.  
\bibitem{sdp}
  M. Haugh, G. Iyengar, I. Song, 
  \emph{A Generalized Risk Budgeting Approach to Portfolio Construction},
  2015.
\bibitem{newton}  
  B. Chaves, C. Hsu, F. Li, O.Shakernia,
  \emph{Efficient Algorithms for Computing Risk Parity Portfolio Weights},
  2012.
 \bibitem{nesterov}
  Y. Nesterov,
  \emph{Introductory Lectures on Convex Programming}, Chap. 4,
  1998.
\bibitem{spinu}
  F. Spinu,
  \emph{Newton's method for the risk parity problem},
  2013.
\bibitem{ccd}
  T. Griveau-Billion, J. Richard, T. Roncalli,
  \emph{A Fast Algorithm for Computing High-dimensional Risk Parity Portfolios},
  2013.
\bibitem{scrip}
  Y. Feng, D. P. Palomar,
  \emph{SCRIP: Successive Convex Optimization Methods for Risk Parity Portfolio Design},
  2015.
\bibitem{genetic}
   R. Hochreiter,
  \emph{An Evolutionary Optimization Approach to Risk Parity Portfolio Selection},
  2015.
  
\bibitem{erb}
  F. Cesarone, F. Tardella,		
  \emph{Equal Risk Bounding is better than Risk Parity for portfolio selection},
  2015.
  
\bibitem{snopt}
	P. E. Gill, W. Murray, M. A. Sanders,
	\emph{SNOPT: An SQP Algorithm for large-scale constrained optimization}, 
	2005.
  
\end{thebibliography}
\end{document}